---
title: Chess Clean
author: Benjamin Labrecque
output: html_document
---

Resource: https://www.youtube.com/watch?v=sV5lwAJ7vnQ&ab_channel=RProgramming101

```{r results='hide'}
library(tidyverse)
```

## Load the data
```{r}
players <- read.csv(Sys.getenv("DATA_PLAYERS"))
```

## Get a glimpse of the data

This shows us the data types of each column and can be useful for knowing if we want
to turn a column from `chr` to a factor column.
```{r}
glimpse(players)
```

What are all possible values for title?
```{r}
unique(players$title)
```

## Convert character column to factor column

```{r}
players$title <- as.factor(players$title)
class(players$title)
```

Set levels in correct order.
```{r}
players$title <- factor((players$title), levels = c("GM", "IM", "FM", "CM", "WGM", "WIM", "WFM", "WCM", "FI", "NI", "DI", "Not ranked"))
levels(players$title)
```

## Selecting columns

```{r}
players %>% 
    select(title, yob, ends_with("der")) %>% 
    names()
```

```{r}
players %>% 
    select(name, title, yob) %>% 
    filter(title %in% c("GM", "IM") & yob > 2005) 
```

## Missing Data

Removing missing data -- this should NOT be the go-to solution. Why is some data missing?
```{r}
mean(players$yob, na.rm = TRUE)

players %>% 
    head()

players %>% 
    na.omit() %>% 
    head()
```

Get all rows with missing values (\<NA\>)
```{r}
players %>% 
    filter(!complete.cases(.)) %>% 
    head()

players %>% 
    filter(!complete.cases(.)) %>% 
    count() # Num rows with at least one missing value

players %>% 
    filter(complete.cases(.)) %>% 
    count() # Num rows with no missing values

players %>% 
    filter(!complete.cases(.)) %>% 
    drop_na(title) %>% 
    head()
```

This shows us that all the missing values are in the "title" column, since
no rows are left once we drop the NA rows from the "title" column.
```{r}
players %>% 
    filter(!complete.cases(.)) %>% 
    drop_na(title) %>% 
    head()
```

Mutate data to replace NA values. If the column is a factor, new values must be in levels.
```{r}
players %>% 
    filter(!complete.cases(.)) %>% 
    mutate(title = replace_na(title, "Not ranked")) %>% 
    head()
```

## Duplicates
```{r}
Names <- c("Peter", "Bob", "Ana", "Peter")
Ages <- c(22, 12, 17, 22)

friends <- data.frame(Names, Ages)
duplicated(friends)
friends[!duplicated(friends),]

friends %>% 
    distinct() # tidyverse way
```

## Recoding variables
```{r}
players %>% 
    select(name, gender) %>%
    mutate(gender_coded = recode(gender, 
                                "M" = 1,
                                "F" = 2)) %>%
    head()
```
